Assimilating GOES-16

### 2021-11-23 analysis/chapters/DA_goes16.Rmd
analisis preliminar ABI

@@

@@

@@

files_ana <- Sys.glob("~/datosmunin3/EXP/prueba_abi/ANA/20181121000000_*/analysis.mem001")

  meta <- unglue(f, "/home/paola.corrales/datosmunin3/EXP/prueba_abi/ANA/20181121000000_{exp}/analysis.mem001")

  ReadNetCDF("~/datosmunin3/EXP/prueba_abi/ANA/20181120180000/analysis.mem001", 
                  vars = c(T_ana = "T", lon = "XLONG", lat = "XLAT")) %>% 
  .[, c("x", "y") := wrf_project(lon, lat)] %>% 
    .[, run := meta[[1]][["exp"]]]

gue <- ReadNetCDF("~/datosmunin3/EXP/prueba_abi/GUESS/20181121000000/wrfarw.mem001", 
                  vars = c(T_guess = "T", lon = "XLONG", lat = "XLAT")) %>% 
  .[, c("x", "y") := wrf_project(lon, lat)]

ana[gue, on = .NATURAL] %>% 
  .[, ":="(level = fcase(bottom_top <= 10, "PBL",
                         bottom_top %between% c(11, 15), "low",
                         bottom_top %between% c(16, 20), "middle",
                         bottom_top %between% c(21, 26), "upper",
                         bottom_top %between% c(27, 36), "the rest"))] %>% 
    .[, a_g := T_ana - T_guess] %>%
  .[level %in% c("low", "middle", "upper") & run %in% c("all", "ch08")] %>% 
  ggplot(aes(x, y)) +
  geom_contour_fill(aes(z = a_g, fill = ..level..), proj = norargentina_lambert) +
  scale_fill_divergent_discretised() +
  geom_mapa() +
  labs(x = NULL, y = NULL, fill = "A-B") +
  facet_grid(run~level) +
  theme_minimal()

ana[gue, on = .NATURAL] %>% 
  .[, ":="(level = fcase(bottom_top <= 10, "PBL",
                         bottom_top %between% c(11, 15), "low",
                         bottom_top %between% c(16, 20), "middle",
                         bottom_top %between% c(21, 26), "upper",
                         bottom_top %between% c(27, 36), "the rest"))] %>% 
    .[, a_g := T_ana - T_guess] %>%
  .[bottom_top == 1] %>% 
  dcast(x + y ~ run, value.var = "a_g") %>% 
  ggplot(aes(x, y)) +
  geom_contour_fill(aes(z = all - noabi, fill = ..level..), proj = norargentina_lambert) +
  scale_fill_divergent_discretised() +
  geom_mapa()

@@

```{r}
ana <- ReadNetCDF("~/datosmunin3/EXP/prueba_abi/ANA/20181120180000/analysis.mem001", 
                  vars = c(T_ana = "QVAPOR", lon = "XLONG", lat = "XLAT")) %>% 
  .[, c("x", "y") := wrf_project(lon, lat)] 
gue <- ReadNetCDF("~/datosmunin3/EXP/prueba_abi/ANA/analysis.mem001",
# gue <- ReadNetCDF("~/datosmunin3/EXP/prueba_abi/GUESS/20181120180000/wrfarw.mem001", 
                  vars = c(T_guess = "QVAPOR", lon = "XLONG", lat = "XLAT")) %>% 
  .[, c("x", "y") := wrf_project(lon, lat)]

ana[gue, on = .NATURAL] %>% 
  .[, a_g := T_ana - T_guess] %>%
  .[bottom_top %in% c(2:7)] %>% 
  ggplot(aes(x, y)) +
  geom_contour_fill(aes(z = a_g, fill = ..level..), proj = norargentina_lambert) +
  scale_fill_divergent_discretised() +
  geom_mapa() +
  facet_wrap(~bottom_top) +
  theme_minimal()

```
```{r}
ana[gue, on = .NATURAL] %>% 
  .[, a_g := T_ana - T_guess] %>%
  .[bottom_top %in% c(10:19)] %>% 
  ggplot(aes(x, y)) +
  geom_contour_fill(aes(z = a_g, fill = ..level..), proj = norargentina_lambert) +
  scale_fill_divergent_discretised() +
  geom_mapa() +
  facet_wrap(~bottom_top) +
  theme_minimal()
```
```{r}
ana[gue, on = .NATURAL] %>% 
  .[, ":="(level = fcase(bottom_top <= 10, "PBL",
                         bottom_top %between% c(11, 15), "low",
                         bottom_top %between% c(16, 20), "middle",
                         bottom_top %between% c(21, 26), "upper",
                         bottom_top %between% c(27, 36), "the rest"))] %>% 
    .[, a_g := T_ana - T_guess] %>%
  .[level %in% c("low", "middle", "upper")] %>% 
  ggplot(aes(x, y)) +
  geom_contour_fill(aes(z = a_g, fill = ..level..), proj = norargentina_lambert) +
  scale_fill_divergent_discretised() +
  geom_mapa() +
  labs(x = NULL, y = NULL, fill = "A-B") +
  facet_wrap(~level) +
  theme_minimal()

@@

```{r}
ana[gue, on = .NATURAL] %>% 
  .[, a_g := T_ana - T_guess] %>%
   .[bottom_top > 10 & lon %between% c(-55, -50)] %>% 
  .[, .(mean_lat = mean(a_g)), by = .(lat, bottom_top)] %>% 
  ggplot(aes(lat, bottom_top)) +
  geom_contour_fill(aes(z = mean_lat, fill = stat(level))) +
  scale_fill_divergent_discretised()
```

>>>>

```{r eval=FALSE, include=FALSE}

source("~/mesoda/analysis/scripts/ejemplo_focus.R")

@@

```{r}
diag %>% 
  .[channel %in% c(8:10)] %>% 
  .[, date := format(date, "%m/%d %H UTC")] %>%
  ggplot(aes(ConvertLongitude(lon), lat)) +
  geom_point(aes(color = factor(qc)), size = 0.9) +
  geom_mapa() +
  scale_color_manual(values = c("darkorange","grey40", "grey60", "grey80"),
                     labels = c("0" = "pass qc", 
                                "3" = "gross check",
                                "5" = "surface check",
                                "50" = "std dev")) +
  guides(color = guide_legend(override.aes = list(size = 2))) +
  facet_grid(channel~date) +
  labs(subtitle = "Assimilated obs", x =  NULL, y = NULL, color = "QC") +
  theme_minimal() +
  theme(legend.position = "bottom")
```

@@

```{r}
files_ana <- Sys.glob("~/datosmunin3/EXP/prueba_abi/ANA/*00/analysis.mem001")

coord <- ReadNetCDF(files_ana[1], vars = c(lon = "XLONG", lat = "XLAT"), subset = list(bottom_top = 1))

levs <- c(200, 300, 400, 500, 600, 700, 800, 900, 1000)

  meta <- unglue(f, "/home/paola.corrales/datosmunin3/EXP/prueba_abi/ANA/{date}/analysis.mem001")

  ReadNetCDF(f, vars = c(t_ana = "T", p = "P", "PB", q_ana = "QVAPOR",
                         lon = "XLONG", lat = "XLAT")) %>% 
    .[, ":="(t_ana = tk(t_ana, p + PB, T_BASE = 290) - 273.15,
             p = p + PB,
             PB = NULL)] %>% 
    .[, date := ymd_hms(meta[[1]][["date"]])] %>% 
    .[, .(lev = levs,
          t_ana = approx(p, t_ana, xout = levs*100)$y,
          q_ana = approx(p, q_ana, xout = levs*100)$y),
      by = .(date, south_north, west_east)] %>%
    coord[., on = .NATURAL] %>% 
    .[, c("x", "y") := wrf_project(lon, lat)] %>% 
    .[]

files_gue <- Sys.glob("~/datosmunin3/EXP/prueba_abi/GUESS/*00/wrfarw.mem001")

gue <- map(files_gue, function(f) {

  meta <- unglue(f, "/home/paola.corrales/datosmunin3/EXP/prueba_abi/GUESS/{date}/wrfarw.mem001")

  ReadNetCDF(f, vars = c(t_gue = "T", p = "P", "PB", q_gue = "QVAPOR",
                         lon = "XLONG", lat = "XLAT")) %>% 
    .[, ":="(t_gue = tk(t_gue, p + PB, T_BASE = 290) - 273.15,
             p = p + PB,
             PB = NULL)] %>% 
    .[, date := ymd_hms(meta[[1]][["date"]])] %>% 
    .[, .(lev = levs,
          t_gue = approx(p, t_gue, xout = levs*100)$y,
          q_gue = approx(p, q_gue, xout = levs*100)$y),
      by = .(date, south_north, west_east)] %>%
    coord[., on = .NATURAL] %>% 
    .[, c("x", "y") := wrf_project(lon, lat)] %>% 
    .[]

```{r}
dates <- unique(ana$date)

for (i in 1:length(dates)) {

  ana[gue, on = .NATURAL] %>% 
  .[date == dates[i]] %>% 
  .[, date := format(date, "%m/%d %H UTC")] %>%
  .[, a_g := t_ana - t_gue] %>%
  .[lev %in% c(300, 500, 700)] %>% 
  ggplot(aes(x, y)) +
  geom_contour_fill(aes(z = a_g, fill = ..level..), proj = norargentina_lambert) +
  scale_fill_divergent_discretised(trans = mesoda::symlog_trans()) +
  geom_mapa() +
  labs(x = NULL, y = NULL, fill = "A-B",
       title = "Temperature",
       subtitle = as.character(dates[i])) +
  facet_grid(date~lev) +
  theme_minimal()

  ggsave(paste0("t_ab_", format(dates[i], "%Y-%m-%d_%H"), ".png"), height = 3)

  ana[gue, on = .NATURAL] %>% 
  .[date == dates[i]] %>% 
  .[, date := format(date, "%m/%d %H UTC")] %>%
  .[, a_g := q_ana - q_gue] %>%
  .[lev %in% c(300, 500, 700)] %>% 
  ggplot(aes(x, y)) +
  geom_contour_fill(aes(z = a_g*1000, fill = ..level..), proj = norargentina_lambert) +
  scale_fill_divergent_discretised(trans = mesoda::symlog_trans()) +
  geom_mapa() +
  labs(x = NULL, y = NULL, fill = "A-B",
       title = "Humidity",
       subtitle = as.character(dates[i])) +
  facet_grid(date~lev) +
  theme_minimal()

  ggsave(paste0("q_ab_", format(dates[i], "%Y-%m-%d_%H"), ".png"), height = 3)

}
```

@@

for (i in 1:length(levs)) {

  ana[gue, on = .NATURAL] %>% 
  .[lev == levs[i]] %>% 
  .[, date := format(date, "%m/%d %H UTC")] %>%
  .[, a_g := t_ana - t_gue] %>%
  ggplot(aes(x, y)) +
  geom_contour_fill(aes(z = a_g, fill = ..level..), proj = norargentina_lambert) +
  scale_fill_divergent_discretised() +
  geom_mapa() +
  labs(x = NULL, y = NULL, fill = "A-B",
       title = "Temperature",
       subtitle = paste(as.character(levs[i]), "hPa")) +
  facet_wrap(~date, ncol = 4) +
  theme_minimal()

  ggsave(paste0("t_ab_", levs[i], ".png"), height = 4)

  ana[gue, on = .NATURAL] %>% 
  .[lev == levs[i]] %>% 
  .[, date := format(date, "%m/%d %H UTC")] %>%
  .[, a_g := q_ana - q_gue] %>%
  ggplot(aes(x, y)) +
  geom_contour_fill(aes(z = a_g*1000, fill = ..level..), proj = norargentina_lambert) +
  scale_fill_divergent_discretised() +
  geom_mapa() +
  labs(x = NULL, y = NULL, fill = "A-B",
       title = "Humidity",
       subtitle = paste(as.character(levs[i]), "hPa")) +
   facet_wrap(~date, ncol = 4) +
  theme_minimal()

  ggsave(paste0("q_ab_", levs[i], ".png"), height = 4)

```
```{r}
files <- Sys.glob("~/datosmunin3/EXP/prueba_abi/ANA/*00/diagfiles/asim_conv*.ensmean")

read_diag_conv(files, "EG1") %>% 
  .[type %in% c(120)] %>% 
  unique(., by = c("stationID", "lat", "lon", "date")) %>% 
  .[] %>% 
  ggplot(aes(ConvertLongitude(lon), lat)) + 
  geom_point() +
  geom_mapa() +
  labs(x = NULL, y = NULL, 
       title = "Soundings",
       subtitle = "2018-11-21 00 UTC")

@@

```{r eval=FALSE, include=FALSE}
ana <- ReadNetCDF("~/datosmunin3/EXP/prueba_abi/ANA/20181120180000/analysis.mem001", 
                  vars = c(T_ana = "QVAPOR", lon = "XLONG", lat = "XLAT")) %>% 
  .[, c("x", "y") := wrf_project(lon, lat)]

gue <- ReadNetCDF("~/datosmunin3/EXP/prueba_abi/GUESS/20181120180000/wrfarw.mem001",
                  vars = c(T_guess = "QVAPOR", lon = "XLONG", lat = "XLAT")) %>% 
  .[, c("x", "y") := wrf_project(lon, lat)]

ana[gue, on = .NATURAL] %>% 
  .[, a_g := T_ana - T_guess] %>%
  .[bottom_top %in% c(2:7)] %>% 
  ggplot(aes(x, y)) +
  geom_contour_fill(aes(z = a_g, fill = ..level..), proj = norargentina_lambert) +
  # scale_fill_divergent_discretised() +
  geom_mapa() +
  facet_wrap(~bottom_top) +
  theme_minimal()

```{r eval=FALSE, include=FALSE}
ana[gue, on = .NATURAL] %>% 
  .[, a_g := T_ana - T_guess] %>%
  .[bottom_top %in% c(10:19)] %>% 
  ggplot(aes(x, y)) +
  geom_contour_fill(aes(z = a_g, fill = ..level..), proj = norargentina_lambert) +
  scale_fill_divergent_discretised() +
  geom_mapa() +
  facet_wrap(~bottom_top) +
  theme_minimal()
```
```{r eval=FALSE, include=FALSE}
ana[gue, on = .NATURAL] %>% 
  .[, ":="(level = fcase(bottom_top <= 10, "PBL",
                         bottom_top %between% c(11, 15), "low",
                         bottom_top %between% c(16, 20), "middle",
                         bottom_top %between% c(21, 26), "upper",
                         bottom_top %between% c(27, 36), "the rest"))] %>% 
  .[, a_g := T_ana - T_guess] %>%
  .[level %in% c("low", "middle", "upper")] %>% 
  ggplot(aes(x, y)) +
  geom_contour_fill(aes(z = a_g, fill = ..level..), proj = norargentina_lambert) +
  scale_fill_divergent_discretised() +
  geom_mapa() +
  labs(x = NULL, y = NULL, fill = "A-B") +
  facet_wrap(~level) +
  theme_minimal()

@@

```{r eval=FALSE, include=FALSE}
ana[gue, on = .NATURAL] %>% 
  .[, a_g := T_ana - T_guess] %>%
  .[bottom_top > 10 & lon %between% c(-55, -50)] %>% 
  .[, .(mean_lat = mean(a_g)), by = .(lat, bottom_top)] %>% 
  ggplot(aes(lat, bottom_top)) +
  geom_contour_fill(aes(z = mean_lat, fill = stat(level))) +
  scale_fill_divergent_discretised()
```
